---
title: "ENMAC scenario"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tmap)
tmap_mode("view")
```

The National targets for cycling uptake were set to:

-   4% of all trips should be made by bicycle by 2025
-   10% of all trips should be made by bicycle by 2030

Cycling trips should replace car trips directly.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
od_sample = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/od_sample.Rds"))
od_sample %>% as.data.frame %>% head(10) %>% knitr::kable()
```


```{r message=FALSE, warning=FALSE, include=FALSE}
topdl = 300 #change here the threshold
```

Top `r topdl` desire lines with most trips. Line width for car trips.

```{r echo=FALSE, message=FALSE, warning=FALSE}
od_all = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/TRIPSmode_freguesias.Rds"))
zones = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/FREGUESIASgeo.Rds"))
# od_all_sf = od::od_to_sf(od_all, zones)

od_all_sf_top = od_all %>% 
  filter(DICOFREor11 != DICOFREde11) %>% 
  top_n(n = topdl, wt = Total)

tm_shape(od_all_sf_top) +
   tm_lines(col = "Total", palette = "viridis", lwd = "Car", scale = 9)
```

### ENMAC: 4% bike, replaced from Car

```{r include=FALSE}
#cycling potential function
od_top_enmac = od_all_sf_top
ENMAC4 = 0.04 # 4%
od_top_enmac$Bikeper = od_top_enmac$Bike / od_top_enmac$Total

od_top_enmac_aside = od_top_enmac %>% filter(Bikeper >= ENMAC4) #don't mess with the ones that already reach target
od_top_enmac = od_top_enmac %>% filter(Bikeper < ENMAC4)

od_top_enmac$new_cyc = ENMAC4 * od_top_enmac$Total - od_top_enmac$Bike
od_top_enmac$new_carT = (od_top_enmac$Car + od_top_enmac$CarP) - od_top_enmac$new_cyc

od_top_enmac_aside$new_cyc = od_top_enmac_aside$Bike
od_top_enmac_aside$new_carT = (od_top_enmac_aside$Car + od_top_enmac_aside$CarP)

od_top_enmac = rbind(od_top_enmac, od_top_enmac_aside)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tm_shape(od_top_enmac) +
  tm_lines(lwd = "new_cyc", scale = 9, col = "new_cyc", palette = "viridis")
```

Example with routing, for top `r topdl` ODs
(fast routes from Cyclestreets.net)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# check route network generation with stplanr
library(stplanr)
routes = sf::read_sf("https://github.com/U-Shift/cyclingpotential-hack/releases/download/2.0.0/routes_fast.geojson")

od_top_enmac$pair = paste0(od_top_enmac$DICOFREor11, od_top_enmac$DICOFREde11)
routes_enmac4 = routes %>% mutate(pair = paste0(DICOFREor11, DICOFREde11)) %>% filter(pair %in% od_top_enmac$pair)
newcyc = od_top_enmac %>% select(pair, new_cyc) %>% sf::st_drop_geometry() %>% mutate(new_cyc = round(new_cyc))
routes_enmac4 = left_join(routes_enmac4, newcyc)

rnet = overline(routes_enmac4, "new_cyc") 
tm_shape(rnet) +
  tm_lines(scale = 15, col = "new_cyc", lwd = "new_cyc", palette = "viridis")

```

With the ENMAC 4% scenario, the cycling potential is **`r round(sum(newcyc$new_cyc)/1000)`k** new cycling trips, transferred from car.


> This is without trips with same O&D. Jittered will bring other results.

### ENMAC: 10% bike, replaced from Car

```{r include=FALSE}
#cycling potential function
od_top_enmac = od_all_sf_top
ENMAC10 = 0.1 # 10%
od_top_enmac$Bikeper = od_top_enmac$Bike / od_top_enmac$Total

od_top_enmac_aside = od_top_enmac %>% filter(Bikeper >= ENMAC10) #don't mess with the ones that already reach target
od_top_enmac = od_top_enmac %>% filter(Bikeper < ENMAC10)

od_top_enmac$new_cyc = ENMAC10 * od_top_enmac$Total - od_top_enmac$Bike
od_top_enmac$new_carT = (od_top_enmac$Car + od_top_enmac$CarP) - od_top_enmac$new_cyc

od_top_enmac_aside$new_cyc = od_top_enmac_aside$Bike
od_top_enmac_aside$new_carT = (od_top_enmac_aside$Car + od_top_enmac_aside$CarP)

od_top_enmac = rbind(od_top_enmac, od_top_enmac_aside)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tm_shape(od_top_enmac) +
  tm_lines(lwd = "new_cyc", scale = 9, col = "new_cyc", palette = "viridis")
```

Example with routing, for top `r topdl` ODs
(fast routes from Cyclestreets.net)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# check route network generation with stplanr
library(stplanr)
routes = sf::read_sf("https://github.com/U-Shift/cyclingpotential-hack/releases/download/2.0.0/routes_fast.geojson")

od_top_enmac$pair = paste0(od_top_enmac$DICOFREor11, od_top_enmac$DICOFREde11)
routes_enmac10 = routes %>% mutate(pair = paste0(DICOFREor11, DICOFREde11)) %>% filter(pair %in% od_top_enmac$pair)
newcyc = od_top_enmac %>% select(pair, new_cyc) %>% sf::st_drop_geometry() %>% mutate(new_cyc = round(new_cyc))
routes_enmac10 = left_join(routes_enmac10, newcyc)

rnet = overline(routes_enmac10, "new_cyc") 
tm_shape(rnet) +
  tm_lines(scale = 15, col = "new_cyc", lwd = "new_cyc", palette = "viridis")

```

And with the ENMAC 10% scenario, the cycling potential is **`r round(sum(newcyc$new_cyc)/1000)`k** new cycling trips, transferred from car.

## Threshold table

```{r echo=FALSE, message=FALSE, warning=FALSE}
scenarios = c("ENMAC 4%", "ENMAC 10%")
thresholds = c(300, 500, 1000, 2000, 5000, nrow(od_all %>% filter(DICOFREor11 != DICOFREde11)))
howmanyt = n

ENMAC4 = 0.04 # 10%
ENMAC10 = 0.1 # 10%

newcyclists4 = numeric()
newcyclists10 = numeric()

# 4%
for(i in 1:length(thresholds)){
  od_all_top = od_all %>% 
    filter(DICOFREor11 != DICOFREde11) %>% 
    top_n(n = thresholds[i], wt = Total)

  od_top_enmac = od_all_top

  od_top_enmac$Bikeper = od_top_enmac$Bike / od_top_enmac$Total

  od_top_enmac_aside = od_top_enmac %>% filter(Bikeper >= ENMAC4) #don't mess with the ones that already reach target
  od_top_enmac = od_top_enmac %>% filter(Bikeper < ENMAC4)

  od_top_enmac$new_cyc = ENMAC4 * od_top_enmac$Total - od_top_enmac$Bike
  od_top_enmac_aside$new_cyc = od_top_enmac_aside$Bike
  od_top_enmac = rbind(od_top_enmac, od_top_enmac_aside)
  
  newcyclists4[i] = round(sum(od_top_enmac$new_cyc))
}

# 10%
for(i in 1:length(thresholds)){
  od_all_top = od_all %>% 
    filter(DICOFREor11 != DICOFREde11) %>% 
    top_n(n = thresholds[i], wt = Total)

  od_top_enmac = od_all_top

  od_top_enmac$Bikeper = od_top_enmac$Bike / od_top_enmac$Total

  od_top_enmac_aside = od_top_enmac %>% filter(Bikeper >= ENMAC10) #don't mess with the ones that already reach target
  od_top_enmac = od_top_enmac %>% filter(Bikeper < ENMAC10)

  od_top_enmac$new_cyc = ENMAC10 * od_top_enmac$Total - od_top_enmac$Bike
  od_top_enmac_aside$new_cyc = od_top_enmac_aside$Bike
  od_top_enmac = rbind(od_top_enmac, od_top_enmac_aside)
  
  newcyclists10[i] = round(sum(od_top_enmac$new_cyc))
}

tablethreshold = data.frame(top = thresholds, enmac4 = newcyclists4, enmac10 = newcyclists10)
names(tablethreshold) = c("top desire lines", scenarios[1], scenarios[2])

tablethreshold[,c(2,3)] = round(tablethreshold[,c(2,3)]/1000,1)

knitr::kable(tablethreshold, caption = "New cycling trips [x1000], by scenario and top desire lines threshold")
# %>%  kableExtra::kable_paper("hover", full_width = F)

```

