---
title: "Validation of the mobility survey for the units of analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
	warning = FALSE
)
```

```{r setup, include = FALSE}
library(dplyr)
library(sf)
library(tmap)
tmap_mode("view")
```

IMOB is representative at zone level, which are groups of _Freguesias_ that have homogeneity in accessibility to public transport, and for each Municipality.

The [methodology of IMOB](https://www.ine.pt/ngt_server/attachfileu.jsp?look_parentBoui=352412255&att_display=n&att_download=y) defines 5 different zone levels, resulting in 49 zones, within the 18 Municipalities.


```{r import data, include=FALSE}
ZONAMENTO = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/ZONAMENTO_imob.Rds"))
FREGUESIASgeo = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/FREGUESIASgeo.Rds"))
MUNICIPIOSgeo = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/MUNICIPIOSgeo.Rds"))
POPFreguesias = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/PopFreguesias.Rds"))
TRIPSmode_freguesias = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/TRIPSmode_freguesias.Rds"))
IMOBresidentes = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/IMOBrespondents.Rds"))
```

```{r echo=FALSE}
FREGUESIASgeo_zones = FREGUESIASgeo %>% left_join(ZONAMENTO %>% select(-2)) %>%
  left_join(POPFreguesias %>% select(-1)) %>% mutate(Zonas = as.integer(Zonas))

tm_shape(FREGUESIASgeo_zones) + tm_polygons(col = "Zonas", palette = "Paired", alpha = 0.8) +
  tm_shape(MUNICIPIOSgeo) + tm_borders(col = "black", lwd = 2)
```

```{r include=FALSE}
TRIPSmode_freguesias_OR = TRIPSmode_freguesias %>% st_drop_geometry() %>%
  mutate(Dicofre = as.character(DICOFREor11)) %>% #origins
  select(3:8,10) %>% group_by(Dicofre) %>% summarise_all(sum)

IMOBvalid_freg = FREGUESIASgeo_zones %>% st_drop_geometry() %>% left_join(TRIPSmode_freguesias_OR) #118
IMOBvalid_zones = IMOBvalid_freg %>% select(-c(1,4)) %>%
  group_by(Concelho, DTMN, Zonas, DTMN_Zona) %>% summarise_all(sum) #49
# sum(IMOBvalid_freg$Total) #5.3M
# sum(IMOBvalid_zones$Total) #5.3M
sum(IMOBvalid_freg$Populacao) #2.8M
sum(IMOBvalid_zones$Populacao) #2.8M
```
According to Census 2021, the population in the Lisbon Metropolitan Area (LMA) was `r round(sum(IMOBvalid_freg$Populacao)/1000)`k, and the number of trips extrapolated from the IMOB 2017 was `r round(sum(IMOBvalid_freg$Total)/1000)`k.

```{r eval=FALSE, include=FALSE, echo=TRUE}
#this was made based on trips, not population
IMOBvalid_freg = IMOBvalid_freg %>% mutate(TRIPS_pop = Total/Populacao, Car_pop = (Car + CarP)/Populacao)
IMOBvalid_zones = IMOBvalid_zones %>% mutate(TRIPS_pop = Total/Populacao, Car_pop = (Car + CarP)/Populacao)

summary(IMOBvalid_freg$TRIPS_pop)
summary(IMOBvalid_zones$TRIPS_pop)

summary(IMOBvalid_freg$Car_pop)
summary(IMOBvalid_zones$Car_pop)


modelfreg = lm(data= IMOBvalid_freg, Total ~ Populacao)
plot(IMOBvalid_freg$Populacao, IMOBvalid_freg$Total) + abline(modelfreg, col = "red")
plot(IMOBvalid_freg$TRIPS_pop) + abline(h = mean(IMOBvalid_freg$TRIPS_pop), col = "red")

modelzones = lm(data= IMOBvalid_zones, Total ~ Populacao)
plot(IMOBvalid_zones$Populacao, IMOBvalid_zones$Total) + abline(modelzones, col = "blue")
plot(IMOBvalid_zones$TRIPS_pop) + abline(h = mean(IMOBvalid_zones$TRIPS_pop), col = "blue")

plot(modelfreg)
plot(modelzones)

#isto é com viagens, mas a metodologia do IMOB tem como referência a população


```

## Validation

We want to check in what extent can we rely on the mobility survey (IMOB), disagregated by _Freguesia_.

The error accepted by IMOB methodology (p.12) is as follows:

```{r echo=FALSE}
tabela_erro = data.frame(popupacao_municipio = c("< 50k", "50-100k", "100-200k", ">= 200k"),
                         error = c("5 %", "4 %", "3 %", "2 %"))
knitr::kable(tabela_erro)
```

And it is computed as:

$$ \varepsilon = \sqrt{ \frac{z ^{2}_{1- \alpha/2} \cdot \widehat{P} ( 1 - \widehat{P} )} {n_{inq}} } $$

With $\varepsilon$ as the error, z=1.96 for a significance level of $\alpha$ as 5%, and $n_{inq}$ as the number of conducted surveys in each area.

```{r include=FALSE}
# filtrar nos microdados o número de inquéritos por zona/municipio

municipios = unique(FREGUESIASgeo_zones$DTMN)
POPMunicipios = POPFreguesias %>% mutate(DTCC = substr(Dicofre, 1, 4)) %>%
  group_by(DTCC) %>% summarise(Populacao = sum(Populacao)) 

POPZona = ZONAMENTO %>% left_join(POPFreguesias) %>% group_by(DTMN_Zona) %>% summarise(PopZona = sum(Populacao))

Pzona = IMOBresidentes %>% filter(DTCC_de %in% municipios) %>%
  mutate(zonamun = paste(DTCC_de, Zona_de, sep="_")) %>%
  left_join(POPFreguesias) %>% 
  group_by(zonamun, DTCC_de) %>% summarize(inquiridos = sum(inquiridos)) %>%
  left_join(POPMunicipios, by=c("DTCC_de" = "DTCC")) %>% 
  left_join(POPZona, by=c("zonamun" = "DTMN_Zona"))

Pzona$classe = if_else(Pzona$Populacao >=200000, 2, ifelse(
  Pzona$Populacao > 100000 & Pzona$Populacao < 200000 , 3,ifelse(
  Pzona$Populacao > 50000 & Pzona$Populacao < 100000 , 4, 5)))

#Proportion
Pzona = Pzona %>% mutate(percent = inquiridos/PopZona)

summary(Pzona$percent)
hist(Pzona$percent)

plot(Pzona$Populacao, Pzona$percent)


#error
Pzona$error = sqrt(1.96^2 * Pzona$percent * (1- Pzona$percent) / Pzona$inquiridos)
summary(Pzona$error)

#inside parameters of validation?
Pzona$confirm = ifelse(Pzona$error*100 < Pzona$classe, "Inside", "Outide")
```


All the 49 zones present errors bellow the acceptable limits defied by IMOB. 

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Zones with error within the parameters defined:
table(Pzona$confirm)
```


```{r include=FALSE}

# filtrar nos microdados o número de inquéritos por freguesia

Pfreg = IMOBresidentes %>% filter(DTCC_de %in% municipios) %>%
  select(Dicofre, inquiridos) %>%
  group_by(Dicofre) %>% summarize(inquiridos = sum(inquiridos)) %>% ungroup() %>% 
  left_join(ZONAMENTO) %>% select(DTMN, DTMN_Zona, Dicofre, inquiridos) %>% 
  left_join(POPFreguesias, by="Dicofre") %>% 
  left_join(POPMunicipios, by=c("DTMN" = "DTCC"))

names(Pfreg)[6] = "PopFreg"
names(Pfreg)[9] = "PopMun"


Pfreg$classe = if_else(Pfreg$PopMun >=200000, 2, ifelse(
  Pfreg$PopMun > 100000 & Pfreg$PopMun < 200000 , 3,ifelse(
  Pfreg$PopMun > 50000 & Pfreg$PopMun < 100000 , 4, 5)))

#Proporcao
Pfreg = Pfreg %>% mutate(percent = inquiridos/PopFreg)

summary(Pfreg$percent)
hist(Pfreg$percent)

plot(Pfreg$PopFreg, Pfreg$percent)


#erro
Pfreg$error = sqrt(1.96^2 * Pfreg$percent * (1- Pfreg$percent) / Pfreg$inquiridos)
summary(Pfreg$error)

#dentro dos parámetros?
Pfreg$valid = ifelse(Pfreg$error*100 < Pfreg$classe, "Yes", "No")
table(Pfreg$valid)
# Nao Sim 
#   7 111 

#identificar as freguesias com +25% (3º quartil)
Pfreg$confianca = ifelse(Pfreg$error*100 > Pfreg$classe, "Above acceptable", ifelse(
  Pfreg$error > summary(Pfreg$error)[5], ">1.8%", ifelse(
  Pfreg$error > summary(Pfreg$error)[3], "1.3-1.8%", ifelse(
  Pfreg$error > summary(Pfreg$error)[2], "1.0-1.3%", "<1.0%"))))

Pfreg$margem_erro = factor(Pfreg$confianca, levels = c("<1.0%", "1.0-1.3%", "1.3-1.8%", ">1.8%", "Above acceptable"))
table(Pfreg$margem_erro)
#?
```


### Error margin for each _Freguesia_

```{r echo=FALSE}
FREGUESIASgeo_zones_error = FREGUESIASgeo_zones %>%
  left_join(Pfreg %>% select(Dicofre, DTMN_Zona, error, valid, confianca, margem_erro)) %>%
  mutate(Error = round(error*100,2))

tm_shape(FREGUESIASgeo_zones_error) + tm_polygons(col = "valid", palette = "Paired")


tm_shape(FREGUESIASgeo_zones_error) +
  tm_polygons(col = "margem_erro",
              title.col = "Error margin",
              popup.vars = c("Dicofre" = "Dicofre",
                   "Concelho" = "Concelho",
                   "Freguesia" = "Name",
                  "Error margin" = "Error"),
           id = "Name") 
# + tm_layout(title= 'Error margin for each Freguesia',  title.position = c('right', 'top'))
  
```

Although there are `r nrow(Pfreg %>% filter(valid=="No"))` _Freguesias_ (in `r nrow(Pfreg)`) with a higher error margin than the accepted levels (Table above), the highest error is of `r round(max(Pfreg$error)*100,2)`%.

```{r echo=FALSE}
tabela = Pfreg %>% select(Dicofre, Freguesia, PopFreg, error, margem_erro, DTMN)
tabela = tabela %>% left_join(FREGUESIASgeo_zones %>% st_drop_geometry() %>% select(DTMN, Concelho)) %>% unique()
tabela = tabela %>% select(Dicofre, Concelho, Freguesia, PopFreg, error, margem_erro) %>% mutate(error = round(error*100,2)) %>% arrange(Concelho, Freguesia)
  
knitr::kable(tabela)
```


## Limitations

-   This comparison was made using Census 2021 population data, and not Census 2011 data (used by IMOB).

-   Regarding Census 2021, the population age is not available yet, in order to filter the ones between 6-84 yrs, as defined as "Active Population".

-   We don't know how many survey respondents were by _Freguesia_. From the Mobility Survey database, we can only figure out their home Municipality and Zone. For this purpose, we considered the _Freguesia_ corresponding with the destination of their stated trip purpose "Return home", discarding 1.28% of all survey respondents.

