---
title: "Prototype"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biclar)
```

This includes some statistics about each Municipality, and the Metro Region.

The maps and statistics include: Baseline Scenario, ENMAC 4%, and ENMAC 10%.

### Separated statistics and maps

-   [**Área Metropolitana de Lisboa**](articles/region.html)
-   [Alcochete](articles/alcochete.html)
-   [Almada](articles/almada.html)
-   [Amadora](articles/amadora.html)
-   [Barreiro](articles/barreiro.html)
-   [Cascais](articles/cascais.html)
-   [Lisboa](articles/lisboa.html)
-   [Loures](articles/loures.html)
-   [Mafra](articles/mafra.html)
-   [Moita](articles/moita.html)
-   [Montijo](articles/motijo.html)
-   [Odivelas](articles/odivelas.html)
-   [Oeiras](articles/oeiras.html)
-   [Palmela](articles/palmela.html)
-   [Seixal](articles/seixal.html)
-   [Sesimbra](articles/sesimbra.html)
-   [Setubal](articles/setubal.html)
-   [Sintra](articles/sintra.html))
-   [Vila Franca de Xira](articles/vfxira.html)


