---
title: "biclaR"
output:
  html_document:
    theme:
      bg: "#FFFFFF"
      fg: "#000000"
      highlight: "#000000"
      primary: "#000000"
      base_font:
        google: "Lexend" #Prompt
      code_font:
        google: "JetBrains Mono"
    css: css/style.css
    includes:
      after_body: img/footer.html
---
  

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=FALSE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)

# has_pub_key = nchar(Sys.getenv("MAPBOX_PUBLIC_TOKEN")) > 5
# knitr::opts_chunk$set(eval = has_pub_key)

library(sf)
library(stringr)
library(dplyr)
library(leaflet)
# library(mapboxapi)
library(DT)
library(htmlwidgets)
```


<h2 style="text-align: center;">Ferramenta de apoio ao planeamento da rede ciclável<br> na Área Metropolitana de Lisboa</h2>

##### Nota: este projeto está em fase _beta_ e a informação deste site deve ser considerada como _work in progress_. 

<!-- Content for landing page-->

```{r pkg, eval=FALSE, include=FALSE}
# TODO this will contain more variables we need, primarily % cycling
# Load counties geometry
counties_raw = readRDS("counties.Rds")
counties_raw = counties_raw %>%
  mutate(county_name = stringr::str_to_title(COUNTY))
# Load pcycle_data at the county level
county_trips = read.csv("county-trips.csv")
# counties_index = county_trips %>%
#   select(county_name, pcycle_baseline)
counties_index = counties_raw %>%
  dplyr::left_join(county_trips)
# counties_index = counties_index %>%
#   dplyr::rename(Name = COUNTY) %>%
#   dplyr::mutate(Name =  stringr::str_to_sentence(Name)) %>%
#   dplyr::mutate(url = paste0(
#     "<a href='",
#     tolower(Name) ,
#     "/' target='_blank'>",
#     Name,
#     "</a>"
#   ))
# counties_index = sf::st_as_sf(counties_index)


municipios_index = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/MUNICIPIOSgeo.Rds")) %>% #it was counties_index -> change to a table with mode statistics!
  # dplyr::mutate(county_name = stringr::str_to_title(Concelho)) %>% #maiusculas
  dplyr::rename(Municipio = Concelho) %>%
  # dplyr::mutate(Municipio = stringr::str_to_sentence(Municipio)) %>% #minusculas
  dplyr::mutate(url = paste0(
  "<a href='", tolower(Municipio) , "/' target='_blank'>",  Municipio,  "</a>"
)) 

municipios_index$url[Municipio == "Vila Franca de Xira"] = paste0("<a href='vfxira/' target='_blank'>",  Municipio,  "</a>")

municipios_index = sf::st_as_sf(municipios_index)


```

# {.tabset}

## Mapa


test content

```{r}
#read data
transit_all = sf::st_read("https://github.com/U-Shift/biclar/releases/download/0.0.1/transit.gpkg", quiet = TRUE)

library(tmap)
tmap_mode("view")

tm_shape(transit_all) +
  tm_sf(col = "mode", size = 0.3)
```


```{r eval=FALSE, include=FALSE}
tippy::tippy_this(elementId = "word", tooltip = "At low zoom, the map shows the % cycle uptake per county; at high zoom, it displays a cycle friendliness measure, with a score between 0 (very unfriendly) and 100 (very friendly). Both measures are shown for each given scenario. See the FAQs for more details")
palette = c(
  '#d73027',
  '#f46d43',
  '#fdae61',
  '#fee090',
  '#ffffbf',
  '#e0f3f8',
  '#abd9e9',
  '#74add1',
  '#4575b4'
)
breaks = c(0, 1, 2, 3, 5, 10, 20, 30, 40, Inf)
pal_baseline = leaflet::colorBin(palette, bins = breaks,
                                 domain = counties_index$pcycle_baseline)
pal_near = leaflet::colorBin(palette, bins = breaks,
                             domain = counties_index$pcycle_near)
pal_climate = leaflet::colorBin(palette, bins = breaks,
                                domain = counties_index$pcycle_climate)
pal_dutch = leaflet::colorBin(palette, bins = breaks,
                              domain = counties_index$pcycle_godutch)
pal_ebike = leaflet::colorBin(palette, bins = breaks,
                              domain = counties_index$pcycle_ebike)
schools_raw = readRDS("schools_all.Rds")
# main map
leaflet::leaflet(counties_index) %>%
  leaflet::addProviderTiles(
    providers$CartoDB.Positron,
    group = "Grey basemap",
    options = leaflet::providerTileOptions(minZoom = 5, maxZoom = 22)
  ) %>%
  leaflet::addPolygons(
    data = counties_index %>% sf::st_transform(crs = 4326),
    color = "black",
    fillColor = "transparent",
    group = "Borders",
    popup = paste0("View county results: ", counties_index$url),
    opacity = 0.7,
    weight = 1.2,
    label = ~ Name,
    labelOptions = leaflet::labelOptions(noHide = F, permanent = TRUE, direction = "auto",sticky = FALSE, textsize = "1.5em")) %>%
  leaflet::addCircles(
    data = schools_raw,
    group = "Schools",
    color = "black",
    opacity = 0.8,
    fillOpacity = 0.3
  ) %>%
  #polygons
  mapboxapi::addMapboxTiles(
    style_id = "cl77hy22i000m14pa2r440vgi",
    username = "nathanaelisamapper",
    group = "Baseline scenario",
    options = leaflet::providerTileOptions(minZoom = 1, maxZoom = 9),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  mapboxapi::addMapboxTiles(
    style_id = "cl77h0i4v001u15o1cp39sk37",
    username = "nathanaelisamapper",
    group = "Near Market",
    options = leaflet::providerTileOptions(minZoom = 1, maxZoom = 9),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  mapboxapi::addMapboxTiles(
    style_id = "cl77hqtom000214qrk6ceqeur",
    username = "nathanaelisamapper",
    group = "Climate Action Plan",
    options = leaflet::providerTileOptions(minZoom = 1, maxZoom = 9),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  mapboxapi::addMapboxTiles(
    style_id = "cl77hvhpx000d14nz0uhmsp40",
    username = "nathanaelisamapper",
    group = "Ebike",
    options = leaflet::providerTileOptions(minZoom = 1, maxZoom = 9),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  mapboxapi::addMapboxTiles(
    style_id = "cl77i0bqh003i16n2cijgn6r1",
    username = "nathanaelisamapper",
    group = "Go Dutch",
    options = leaflet::providerTileOptions(minZoom = 1, maxZoom = 9),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  #rnet
  mapboxapi::addMapboxTiles(
    style_id = "cl66fjj97000a15ru2hhss06g",
    username = "nathanaelisamapper",
    group = "Baseline scenario",
    options = leaflet::providerTileOptions(minZoom = 9, maxZoom = 22),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  mapboxapi::addMapboxTiles(
    style_id = "cl66f8ax6001i14ti630sk73z",
    username = "nathanaelisamapper",
    group = "Climate Action Plan",
    options = leaflet::providerTileOptions(minZoom = 9, maxZoom = 22),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%   mapboxapi::addMapboxTiles(
    style_id = "cl66foarr001d14lmc88wyu42",
    username = "nathanaelisamapper",
    group = "Near Market",
    options = leaflet::providerTileOptions(minZoom = 9, maxZoom = 22),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  )  %>% mapboxapi::addMapboxTiles(
    style_id = "cl66flaic000114t385c8spxn",
    username = "nathanaelisamapper",
    group = "Ebike",
    options = leaflet::providerTileOptions(minZoom = 9, maxZoom = 22),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>% mapboxapi::addMapboxTiles(
    style_id = "cl66fmn9j001g14lkbone3hye",
    username = "nathanaelisamapper",
    group = "Go Dutch",
    options = leaflet::providerTileOptions(minZoom = 9, maxZoom = 22),
    access_token = Sys.getenv("MAPBOX_PUBLIC_TOKEN")
  ) %>%
  leaflet::setView(lng = -8.2,
                   lat = 53.5,
                   zoom = 7) %>%
  leaflet::addLegend(
    "bottomleft",
    colors = c("#882255",  "#CC6677", "#44AA99", "#117733"),
    labels = c("0 to 25", "25 to 50", "50 to 75", "75 to 100"),
    title = "Cycle friendliness",
    group = "Friend",
    opacity = 0.7
  ) %>%
  leaflet::addLegend(
    "bottomleft",
    colors = c(
      '#d73027',
      '#f46d43',
      '#fdae61',
      '#fee090',
      '#ffffbf',
      '#e0f3f8',
      '#abd9e9',
      '#74add1',
      '#4575b4'
    ),
    labels = c(
      "0 to 1",
      "1 to 2",
      "2 to 3",
      "3 to 5",
      "5 to 10",
      "10 to 20",
      "20 to 30",
      "30 to 40",
      "40 to more"
    ),
    title = "Percent cycling",
    group = "Base",
    opacity = 0.7
  ) %>%
  leaflet::addLegend(
    title = "Get Started:",
    group = "Help",
    position = "bottomright",
    colors = c("white",
               "white"),
    labels = c(
      "Zoom in to explore the network",
      "Click on a county for detailed results"
    )
  ) %>%
  addLayersControl(
    baseGroups  = c(
      "Baseline scenario",
      "Near Market",
      "Climate Action Plan",
      "Go Dutch",
      "Ebike"
    ),
    overlayGroups = c("Schools"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  ) %>%
  leaflet::groupOptions(group = "Help", zoomLevels = 7:8)  %>%
    leaflet::groupOptions(group = "Base", zoomLevels = 1:9) %>%
  leaflet::groupOptions(group = "Friend", zoomLevels = 10:20) %>%
  leaflet::hideGroup("Schools")  %>%
  htmlwidgets::onRender("
        function() {
            $('.leaflet-control-layers-base').prepend('<label style=\"text-align:center;font-size:larger;\"><b>Scenarios</b></label>');
        }
    ")
```

**O que mostra este mapa? **  
Por enquanto nada.


<!-- ------------------------------------------------------------------------ -->

## AML

Construir tabela com os dados de populacao, Bike, cyc4, ebike, etc

```{r eval=FALSE, include=FALSE}
# Aim: Build table for homepage
# dn = "popupCycleways/v1"
# built_counties = list.files(dn)
# set.seed(42)
municipios_index %>%
  sf::st_drop_geometry() %>%
  # dplyr::select(url, Population, "% cycling (Baseline)" = pcycle_baseline, "% cycling (Near Market)" = pcycle_near,  "% cycling (Climate Action Plan)" = pcycle_climate,  "% cycling (Go Dutch)" = pcycle_godutch,  "% cycling (Ebike)" = pcycle_ebike) %>%
  # dplyr::rename(Name = url) %>%
  dplyr::arrange(Name) %>%
  # dplyr::mutate_if(is.numeric, round, 1) %>%
  DT::datatable(escape = -2, filter = 'none', options = list(pageLength = 18,
                                                             responsive = TRUE,
                                                             extensions = c('Responsive'))) 
```

ver página com detalhes: [AML](aml/index.html)


<!-- ------------------------------------------------------------------------ -->

## Municípios

Construir tabela com os dados de populacao, Bike, cyc4, ebike, etc, por municipio

```{r}
# Aim: Build table for homepage
# dn = "popupCycleways/v1"
# built_counties = list.files(dn)
# set.seed(42)
municipios_index %>%
  sf::st_drop_geometry() %>%
  dplyr::left_join(HEAT_municipios_PPP_redux %>% filter(CodePath == 1304)) %>% 
  dplyr::select(url, Pop21, Total, Bike, Bike_new, CO2eq10, Economic10) %>%
  dplyr::mutate(Economic10 = Economic10/1000,
                CO2eq10 = as.numeric(gsub(" ","", CO2eq10))) %>% 
  # Population, "% cycling (Baseline)" = pcycle_baseline, "% cycling (Near Market)" = pcycle_near,  "% cycling (Climate Action Plan)" = pcycle_climate,  "% cycling (Go Dutch)" = pcycle_godutch,  "% cycling (Ebike)" = pcycle_ebike) %>%
  dplyr::rename(Municipio = url) %>%
  dplyr::arrange(Municipio) %>%
  # dplyr::mutate_if(is.numeric, round, 1) %>%
  DT::datatable(escape = -1, #para colocar os links, a contar com os rownames
                filter = 'none',
                rownames = FALSE,
                colnames = c("Município", "População 2021", "Viagens /dia", "Viagens Bicicleta", "Novas Viagens Bicicleta 4%",
                             "CO2eq evitado em 10 anos (ton)", "Benefícios estimados em 10 anos (milhares €)"),
                options = list(dom = "t",
                               pageLength = 18,
                               responsive = TRUE,
                              extensions = c('Responsive'))) %>% 
  DT::formatRound(c(2:7),0, mark = " ")
```



<!-- ------------------------------------------------------------------------ -->

## Sobre

### Objetivos da ferramenta biclaR

This information can be used to inform decisions about where best to invest in new or improved cycle routes.

### A equipa biclaR

*   Transportes Metropolitanos de Lisboa (coordenação): XX XXX XXXX.

*   Instituto Superior Técnico - Universidade de Lisboa (desenvolvimento): Filipe Moura, Rosa Félix.

*   Institute for Transport Studies - University of Leeds (apoio ao desenvolvimento): Robin Lovelace.

### Agradecimentos

*   Grupo de Trabalho: ...
*   Conselho Consultuvo: ....
*   Matheus Correia, OpenStreetMap Portugal
*   CycleStreets
*   r5r

### Feedback

You can give feedback on the tool in the [online survey](nada).

### Contactos

Enviar um email para...


<!-- ------------------------------------------------------------------------ -->

## FAQ

Short answers to key frequently answered questions (FAQ) are provided below.
See the full FAQ page for more details and answers to questions about how the tool works.

### O que é a ferramenta biclar?

tal tal.

### Qual o objetivo da ferramenta biclar?

tal tal tal.


<!-- ------------------------------------------------------------------------ -->



```{r eval=FALSE, include=FALSE}
# ?
setTimeout(function(){
  $('.loader-bg').fadeToggle();
}, 1000);
```


